---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rd4utils

<!-- badges: start -->
<!-- badges: end -->

rd4utils wraps the [d4utils](https://github.com/38/d4-format) tool for reading &
writing d4 files using R data structures.

## Installation

### Development Version
You can install the development version of rd4utils from [GitHub](https://github.com/snystrom/rd4utils) with:

```{r, eval=F}
if (!requireNamespace("remotes", quietly=TRUE))
  install.packages("remotes")
remotes::install_github("snystrom/rd4utils")
```

```{r}
library(rd4utils)
```

### d4utils

rd4utils relies on a local install of
[d4utils](https://github.com/38/d4-format). You can find binaries at the
[d4utils releases](https://github.com/38/d4-format/releases) page, or if you
have [Rust](https://www.rust-lang.org/tools/install) installed, can compile from
source using `cargo install d4utils`.

rd4utils needs to know the location of `rd4utils` on your local machine. If
installing using cargo, rd4utils should autodetect the binary if installed at
`~/.cargo/bin/d4utils`. Otherwise, you can set the `d4utils` option once per
session like so:

`options("d4utils" = "path/to/d4utils")`

Alternatively, all rd4utils functions contain the `d4utils` argument, which
accepts a path to the `d4utils` binary.

To test whether R detects d4utils, run 
```{r}
d4_install_check()
```

## Examples

 - `d4_view()`
 - `d4_view_genome()`
 - `d4_stat()`
 - `d4_hist()`

# Citation

Please cite the d4 authors if using rd4utils:

Efficient storage and analysis of quantitative genomics data with the Dense
Depth Data Dump (D4) format and d4tools. Hao Hou, Brent S Pedersen, Aaron
Quinlan bioRxiv 2020.10.23.352567; doi: https://doi.org/10.1101/2020.10.23.352567
